---
title: "Final Project R Reference"
author: "Juliana Berube"
date: "12/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# R Reference Guide {.tabset .tabset-pills}


## Loading Data and Packages

To use load data using "require" and "here"

```{r}
require(here)
library (palmerpenguins)

#To load the ginkgo data:

ginkgo = read.csv(here("data", "ginkgo_data_2021.csv"))
```


## Data Structures

### c( )
The function c( ) combines or concatenates its arguments into a vector (a 1-dimensional data structure consisting of 1 or more elements).

* All of the elements must be of the same type.
* I can’t combine character and numeric types in the same call to c( )

Here’s two examples using numeric and character data types:

```{r}
## Create a vector of numbers:
num_vec  = c(1, 4, 8, 9, 13)

## Create a vector of characters:
char_vec = c("a", "fish", "data is cool")
```

I can show the contents of a vector by typing the name of the vector, or using the print() function.

```{r}
## Typing the name of the vector into the console prints the contents
num_vec


## The print() function accomplishes the same task:
print(char_vec)
```

### length( )

"Get or set the length of vectors (including lists) and factors, and of any other R object for which a method has been defined" - R Help

Arguments can either be x (R object) or a value (interger)

```{r}
length(num_vec)
```

In a data frame, the length argument is restricted to the number of columns:
```{r}
length(ginkgo)
```

Can also be used inside loops and other functions to find the length or number of repetitions (come back to this)

### Matrix ( )

An array (2 dimensional) of numbers or characters arranged into rows and columns. 

Formula: 

matrix(data = NA, nrow = 1, ncol = 1, byrow = FALSE,
       dimnames = NULL)
       
```{r}
matrix(1:9, nrow = 3, ncol = 3)
```

The automatic arrangment is columns, then row. 

You can also turn vectors into matrixes. Here is an example of creating multiple matrixes with different lengths from one vector. 

```{r}
my_vec<-c(1,2,3)
mat_1 = matrix(my_vec)
mat_1[3,1]

vec_2<-c("one", "two", "three")
vec_3<-c(1.0, 2.0, 3.0)

mat_2=matrix(my_vec, nrow=2, ncol = 3)
mat_3=matrix(my_vec, nrow=3, ncol = 2)
mat_4=matrix(my_vec, nrow=5, ncol=5)
```
See, mat_4 gets and error because the number of rows and columns is longer than my_vec, which is what each matrix is drawing from to populate. 

You can use length() here too to get the length of a matrix

```{r}
length(mat_2)
```

### data.frame ( )

This function will create a data frame. Data frames store data tables and can contain multiple types of data in field (whereas a matrix has all the same type of data). Each column contains one variable, and each row contains one set of values from each column. A data frame can be created like this:

```{r}
ex_month.data <- data.frame(
   month = c (1:12), 
   month_name = c("Jan","Feb","Mar","Apr","May", "Jun", "July", "Aug", "Sep", "Oct", "Nov", "Dec"),
   sunny_days = c(20,15,19,10,22,25, 28, 21, 17, 16, 18, 14)) 
```

### nrow( ), ncol( ), dim( )

You can also find the number of rows and columns in a data frame like this:

```{r}
nrow(ginkgo)
ncol(ginkgo)
dim(ginkgo)
```

## Subsetting

Lets go through some subsets using our ginkgo data. Some ways to do this is to use $ to subset by column name, [] to subset by name, position, row, or column, and subset()

```{r}
ginkgo$max_depth

#you can assign this to a seperate data frame by naming it: depth=ginkgo$max_depth
#[] subsetting chooses by row, column
#select first row of the ginkgo data:
ginkgo[1,]

#select the element in row 2, column 3:
ginkgo[2,3]

#select the 3rd column of the ginkgo data
ginkgo[,3]

#retrieve all the data for Adelie penguins (in the species column) from the penguins dataset
subset(penguins, species == "Adelie")
```

## Numerical Data Exploration

To view our data we can use functions like summary(), mean(), and sd()

```{r}
#Summary of the entire dataset
summary(ginkgo)

#Summary of a column
summary(ginkgo$max_width)

#mean of a column
mean(ginkgo$max_width)

#sd of a column
sd(ginkgo$max_width)

```

## Graphical Data Exploration

To represent data my favorite way: graphically. 

### Scatterplots
Lets begin with a scatterplot...

```{r}
plot(max_width~max_depth, data = ginkgo,
     xlim=c(5,145),
     ylim=c(5, 130),
     xlab= "Leaf Depth (max)",
     ylab= "Leaf Width (max)",
     main= "Ginkgo Leaf: Width and Depth",
     col= "goldenrod",
     cex= 0.5,
     pch=5)
```
So pretty! 

### Histogram 

Let's do a histogram of penguin data:

```{r}
hist(penguins$flipper_length_mm,
     main="Hist of Penguin Flipper Length",
     xlab="Flipper Length (mm)",
     col="azure3",
     breaks=6)
```

### Boxplots

Moving onto boxplots:

1. A classic boxplot
```{r}
boxplot(ginkgo$petiole_length, 
        main="Ginkgo Petiole Length")


```

2. A conditional boxplot (conditioned on seeds_present) 
```{r}
boxplot(max_depth~seeds_present,data=ginkgo,
        main="Ginkgo Leaf Max Depth by Seeds Present",
        xlab="Seeds Present",
        ylab="Max Depth",
        col=c("darkgoldenrod","darkgoldenrod1"))
```

### Panels

```{r}
par(mfrow=c(2,2))

hist(ginkgo$notch_depth,
     main="Hist of Ginkgo Leaf Notch Depth",
     xlab="Notch Depth",
     col="goldenrod",
     breaks=c(0,15,25,50))
hist(penguins$body_mass_g,
     main="Hist of Penguin Body Mass",
     xlab="Body Mass(g)",
     col="azure3",
     breaks=5)
hist(ginkgo$petiole_length,
     main="Hist of Ginkgo Leaf Petiole Length",
     xlab="Notch Depth",
     col="goldenrod2")
hist(penguins$bill_depth_mm,
     main="Hist of Penguin Bill Depth",
     xlab="Bill Depth (mm)",
     col="azure4",
     breaks=8)
```




## Distribution Functions

These functions calculate the probability of events for different parametric distributions.

For a discrete distribution (binomial), the probability mass of an event is the probability that the event will happen. 

Each functions has a different argument, depending on the specific parameters of the distribution.

### dnorm( ) 

dnorm(x, mean = 0, sd = 1, log = FALSE): the probability density

```{r}
dnorm(-1, mean = 0, sd = 1, log = FALSE)

#Basic plot of normal curve
# Generate a vector of x-values
x = seq(-3, 3, length.out = 1000)
y = dnorm(x)

plot(x, y, main = "Normal PDF", type = "l")
abline(h = 0)
```

### pnorm( )

pnorm(q, mean = 0, sd = 1, lower.tail = TRUE, log.p = FALSE): the cumulative probability density

```{r}
pnorm(-1.96, mean = 0, sd = 1, log = FALSE)
pnorm(20, mean=10.4, sd=2.4)

```


### qnorm( )

qnorm(p, mean = 0, sd = 1, lower.tail = TRUE, log.p = FALSE): the quantile function

```{r}
#find the Z-score of the 99th quantile of the standard normal distribution 
qnorm(.99, mean=0, sd=1)
```

### dbinom( ) 

```{r}
#Probabilty of observing four out of 6 with a 2/3 probabilty of observation
dbinom(4,6, 2/3)

#Probablity of observing 0
dbinom(0, 6, 2/3)
```

### pbinom( )
The p-functions calculate the probability of observing a range of events (cumulative probability)

```{r}
#Probabilty of observing 4 or fewer out of 6 with a 2/3 probabilty of observation
pbinom(4, 6, 2/3)

#4 or greater 
1-pbinom(3, 6, 2/3)
```


### qbinom( )
```{r}
#find the 10th quantile of a binomial distribution with 10 trials and prob of success on each trial = 0.4
qbinom(.10, 10, .4)
```

That's it for now!
